---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# condsurv

This package contains a function for generating conditional survival estimates with associated confidence intervals, and a function for plotting conditional survival curves.

Install the package using

```{r eval = FALSE}
remotes::install_github("zabore/condsurv")
```


# Conditional survival

If $S(t)$ represents the survival function at time $t$, then conditional survival is defined as 

$$S(y|x) = \frac{S(x + y)}{S(x)}$$

where $y$ is the number of additional survival years of interest and $x$ is the number of years a subject has already survived.


## Generating conditional survival estimates

The `conditional_surv_est` function will generate this estimate along with 95\% confidence intervals.

The `lung` dataset from the `survival` package will be used to illustrate.

```{r message = FALSE}
library(survival)
library(dplyr)

# Scale the time variable to be in years rather than days
lung2 <- 
  mutate(
    lung,
    os_yrs = time / 365.25
  )
```

First generate a single conditional survival estimate. This is the conditional survival of surviving to 1 year conditioned on already having survived 6 months. This returns a list, where `cs_est` is the conditional survival estimate, `cs_lci` is the lower bound of the 95\% confidence interval and `cs_uci` is the upper bound of the 95\% confidence interval.

```{r}
library(condsurv)

myfit <- survfit(Surv(os_yrs, status) ~ 1, data = lung2)

conditional_surv_est(
  basekm = myfit,
  t1 = 0.5, 
  t2 = 1
)
```

You can easily use `purrr::map_df` to get a table of estimates for multiple timepoints. For example we could get the conditional survival estimate of surviving to a variety of different time points given that the subject has already survived for 6 months (0.5 years).

```{r}
prob_times <- seq(1, 2.5, 0.5)

purrr::map_df(
  prob_times, 
  ~conditional_surv_est(
    basekm = myfit, 
    t1 = 0.5, 
    t2 = .x) 
  ) %>% 
  mutate(years = prob_times) %>% 
  select(years, everything()) %>% 
  knitr::kable()
```


## Plotting conditional survival curves

To plot the conditional survival curves at baseline, and for those who have survived 6 months, 1 year, 1.5 years, and 2 years, we use the `gg_conditional_surv` function.

```{r fig.width = 6, fig.height = 4}
cond_times <- seq(0, 2, 0.5)

gg_conditional_surv(
  basekm = myfit, 
  at = cond_times,
  main = "Conditional survival in lung data"
  ) 
```




