---
title: "Estimate conditional survival"
author: "Emily C. Zabor"
date: "Last updated: `r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Estimate conditional survival}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(condsurv)
library(dplyr)
library(survival)
```


If $S(t)$ represents the survival function at time $t$, then conditional survival is defined as 

$$S(y|x) = \frac{S(x + y)}{S(x)}$$

where $y$ is the number of additional survival years of interest and $x$ is the number of years a subject has already survived.


## Generating conditional survival estimates

The `conditional_surv_est` function will generate this estimate along with 95\% confidence intervals.

The `lung` dataset from the `survival` package will be used to illustrate.

```{r message = FALSE}
# Scale the time variable to be in years rather than days
lung2 <- 
  mutate(
    lung,
    os_yrs = time / 365.25
  )
```

First generate a single conditional survival estimate. This is the conditional survival of surviving to 1 year conditioned on already having survived 6 months ($0.5$ year). This returns a list, where `cs_est` is the conditional survival estimate, `cs_lci` is the lower bound of the 95\% confidence interval and `cs_uci` is the upper bound of the 95\% confidence interval. 

```{r}
myfit <- survfit(Surv(os_yrs, status) ~ 1, data = lung2)

conditional_surv_est(
  basekm = myfit,
  t1 = 0.5, 
  t2 = 1
)
```

You can easily use `purrr::map_df` to get a table of estimates for multiple timepoints. For example we could get the conditional survival estimate of surviving to a variety of different time points given that the subject has already survived for 6 months (0.5 years).

```{r}
prob_times <- seq(1, 2.5, 0.5)

purrr::map_df(
  prob_times, 
  ~conditional_surv_est(
    basekm = myfit, 
    t1 = 0.5, 
    t2 = .x) 
  ) %>% 
  dplyr::mutate(years = prob_times) %>% 
  dplyr::select(years, everything()) %>% 
  knitr::kable()
```

## A note on confidence interval estimation

The confidence intervals are based on a variation of the log-log transformation, also known as the "exponential" Greenwood formula, where the conditional survival estimate is substituted in for the traditional survival estimate in constructing the confidence interval.

If $\hat{S}(y|x)$ is the estimated conditional survival to $y$ given having already survived to $x$, then 

$$\hat{S}(y|x)^{exp(\pm1.96\sqrt{\hat{L}(y|x)})}$$

where 

$$\hat{L}(y|x)=\frac{1}{\log(\hat{S}(y|x))^2}\sum_{j:x \leq \tau_j \leq y}\frac{d_j}{(r_j-d_j)r_j}$$

and

$\tau_j$ = distinct death time $j$

$d_j$ = number of failures at death time $j$

$r_j$ = number at risk at death time $j$
